---
title: "K-means & K-medoid clustering in product segmentation: ASDS 6303 Final Project"
author: "Submitted by Utkarsh Pant"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
library(ggplot2)
library(GGally)
library(readxl)
library(factoextra)
library(cluster)
```

## Loading the dataset

```{r}
product_data = read_excel('./dataset/sku_data.xlsx')
kable(head(product_data),
      booktabs = TRUE,
      format = "latex",
      caption = "Dataset head") %>% kable_styling(latex_options = "hold_position")
```

```{r}
summary(product_data)
```

```{r}
product_data <- select(product_data, -c("ID"))
```

## Checking correlation

```{r}
library(ggcorrplot)
correlation = cor(product_data)
ggcorrplot(correlation, hc.order = TRUE, type = "lower",
   lab = TRUE)
```

Let's only consider the `Outbound number` and `Total outbound` features in our dataset to perform the clustering, due to high correlation.

```{r}
product_subset <- select(product_data, c("Outbound number", "Total outbound"))
```

## Scaling data

```{r}
product_subset_scaled = scale(product_subset)
```

## K-means clustering

Checking a scree-plot for the ideal number of clusters, we see:

```{r}
fviz_nbclust(product_subset_scaled, kmeans, method = "wss") +
geom_vline(xintercept = 3, linetype = 2)
```

```{r}
set.seed(8093)
model.kmeans <- kmeans(product_subset_scaled, nstart = 20, centers = 3)
print(model.kmeans)
```

## Aggregating cluster characteristics

```{r}
kable(aggregate(product_data, by=list(cluster=model.kmeans$cluster), mean),
      format = "latex",
      booktabs = TRUE) %>% kable_styling(position="center")
```

```{r}
fviz_cluster(model.kmeans, product_subset_scaled)
```

## Silhouette scores

[TODO]

# K-medoid clustering

Visualizing the ideal number of clusters for `pam` (k-medoid clustering) using `fvz_nbclust`:

```{r}
fviz_nbclust(product_subset_scaled, pam, method = "wss") +
  geom_vline(xintercept = 5, linetype = 2)

```

It appears that k-medoid clustering for the same product-data does best with 5 clusters. Performing clustering with 5 clusters:

```{r}
set.seed(8093)
model.kmedoid <- pam(product_subset_scaled, k = 5)
model.kmedoid
```

## Aggregating cluster characteristics

```{r}
kable(aggregate(product_data, by=list(cluster=model.kmedoid$cluster), mean),
      format = "latex",
      booktabs = TRUE) %>% kable_styling(position="center")
```

```{r}
fviz_cluster(model.kmedoid, product_subset_scaled)
```
